PoVaTiM: A Possibilistic Valid-Time Model \\ \vspace*{0.04truein}

### 2012-06-23 ijcis-pos-model/sources/temporaldatabases.tex
more modifications

\begin{itemize}
 \item A set of non-temporal attributes.
	\begin{equation}
	\label{eq:attribute-set}
	A = \left \lbrace A_1, A_2, \ldots, A_n \right \rbrace
	\end{equation}
      The domain for each attribute $A_1, \ldots, A_n$ is $D_1, \ldots, D_n$ respectively. 
\item The primary key for the attributes in $A$ is given by $A_K$:
      \begin{equation}
       \label{eq:primary-key-a}
      A_K \subset A
      \end{equation}
\item A set of two attributes; $S$  and $E$ the attributes for the starting and ending points respectively. $I$ defines the valid time interval for the data. 
\begin{equation}
 \label{eq:attribute-time-interval}
I = \left \lbrace S, E \right \rbrace
\end{equation}
\item Then $R$, the schema for the valid-time relation is:
\begin{equation}
 \label{eq:valid-time-relation}
R = A \cup  I
\end{equation}
\item The primary key for the valid-time relation $R$ is:
\begin{equation}
 \label{eq:valid-time-temporal-pk}
PK = A_K \cup I
\end{equation}

@@

      \item Let $A_k \subset A$ be the set of the non-temporal attributes that are part of the primary key (equation \eqref{eq:primary-key-a}). Then, $a_k$ denotes the values for the non-temporal attributes of the primary key.
	    \begin{equation}
	     \label{eq:all-the-versions} 
	      a_k = t\left[A_K \right]
	    \end{equation}

@@

\begin{samepage}
\vglue13pt
%\begin{table}[htbp]
\tcap{\label{tbl:sample-definitions}Sample database containing the instance $r$ of the valid-time relation $R$.}
%\centerline{\small DATA TYPES}
\vglue-6pt
\centerline{\small\baselineskip=13pt
\begin{tabular}{c c c c c c}\\
\hline
& \textbf{A$_1$}  & \textbf{A$_2$}  & $A_3$ & $S$ & $E$ \\
\hline
$t_1$&$a_{11}$ & $a_{12}$ & $a_{13}$ & $s_1$ & $e_1$ \\
$t_1$ & $a_{21}$ & $a_{22}$ & $a_{23}$ &  $s_2$ & $e_2$ \\
$t_1$ & $a_{11}$ & $a_{12}$ & $a_{31}$ & $s_3$ & $e_3$\\
\hline\\
\end{tabular}
}
\end{samepage}

@@

For example, let us consider the last row in Table \ref{tbl:library-sample}. The value for the time interval is $I = \left(4/4/2012 , UC \right)$. The meaning is that the patient with ID = 3 was in the hospital the  4/4/2012 and he or she is still there. The patient with ID = 3 is current in the relation.

\begin{example}
\label{ex:library-database}
 Consider a hospital database. Each patient is stored in the database with an identifier (ID), the name and the dates when the patient entered and left the hospital. Table \ref{tbl:library-sample} illustrates the first version of the database, after three insertions. In this example, $A = ($ ID, Name $)$ and $I = ($ Arrival, Leave $)$.
\end{example}

\begin{samepage}
\vglue13pt
%\begin{table}[htbp]
\tcap{\label{tbl:library-sample}Sample hospital database}
%\centerline{\small DATA TYPES}
\vglue-6pt
\centerline{\small\baselineskip=13pt
\begin{tabular}{c c c c}\\
\hline
\textbf{ID}  & Name  & Arrival & Leave \\
\hline
3 & John & 15/3/2012 & 30/3/2012 \\
4 & Antoon & 25/3/2012 & 4/4/2012 \\
5 & Daan & 18/3/2012 & 2/4/2012\\
3 & John & 4/4/2012 & UC \\
\hline\\
\end{tabular}
}
\end{samepage}

\begin{definition}
\label{def:current-in-relation}
\emph{Current $\left(a_k \right)$.}
Consider the elements in definition \ref{def:valid-time-relation}. The function Current $\left(a_k \right)$ returns the crisp time interval $i$ if the tuple $t$ with the primary key values given by $a_k$ is currently in the instance $r$ of the relation $R$. The function returns the emtyset otherwise.
% 
% Consider $A=\left(A_1, \ldots, A_n \right), A \in \A$ be the set of attributes that define an entity and $I = \left(S_i, UC \right)$ the set of attribute defining the time interval. Let $R \subseteq \A\  x\  \I$ be a valid-time relation. The function Current $\left(R, A \right)$ returns the crisp time interval $I$ if the entity $A$ is currently in the relation $R$ or the empty set, if the entity $A$ is not currently in the relation. It is defined as follows:

@@

\begin{align}
\label{eq:current-in-relation}
\mbox{Current} \left(a_k\right) &=& \\ 
\begin{cases}
\nonumber
t & \mbox{ if } \exists t \in r : i = \left(s, UC \right)\\
\emptyset & \mbox{ otherwise}
\end{cases}
\end{align}
\end{definition}

For example, the function Current(R, (3,`John')) returns the time interval $I = \left(4/4/2012 , UC  \right)$. Conversely, Current(R, (4,`Antoon')) returns the empty set.

@@

\begin{definition}
\label{def:close-current-version}
\emph{Close-current$\left(r, t\right)$}.
Consider the elements in definition \ref{def:valid-time-relation}. The function Close-current$\left(r, t \right)$ closes any current version $t_k$ of the entity given by $t$ if it exists and add the new version $t$. For the implementation the following variables are defined:
\begin{align}
 \label{eq:variables-current-version}
t_{CUR} = \mbox{ Current } \left(t[A_K] \right)\\
\nonumber
t_{UP} = t_{CUR}; t_{UP}\left[i \right] = \mbox{Close }\left(t_{UP}\left[i \right], t \right)
\nonumber
\end{align}
Then, $t_{CUR}$ is the current version of the entity given by the tuple $t$, and $t_{UP}$ is the updated version of the tuple $t_{CUR}$. In this updated version, the time interval given by $i_{UP}$ is closed with respect to the tuple $t$.

@@

For example, consider the patient with ID = 3 in Table \ref{tbl:library-sample}. The current version arrived on 4/4/2012. The patient left the hospital but, for some reason, the date  was not registered. The patient arrived again at the hospital on 24/4/2012. Then, the Close-current function allows to close the current version of the patient and insert a new version. First, the function CloseR is applied with $I = \left(4/4/2012, UC \right)$ and $J = \left(24/4/2012, UC \right)$. Hence, the value for the time interval $I$ is $ \left(4/4/2012, 23/4/2012 \right)$. Then, the modifications on the value for the time interval $I$ are stored. Finally, a new row with the current values of the patient and the time interval $J$ is stored.
The result of this operation is illustrated on Table \ref{tbl:library-sample-2}.

%It is also possible to append at the beggining by using \eqref{def:close-a-crisp-interval-l} and \eqref{def:first-in-relation}.
%\begin{definition}
%\label{def:append-first-version}
%Consider an object $A$, the relation $R$ and a time interval $J$. Then, the function append-first$\left(R,A,J \right)$ closes any right open version of the object $A$ if it exists and add the new version $\left(A, J \right)$.
%
%\begin{eqnarray}
%%\label{eq:close-current}
%\text{Append-first} \left(R, A, J \right) =\\
%\begin{cases}
%\nonumber
%\big \lbrace R - \left(A, I \right) \cup \left \lbrace \left(A, \mbox{ CloseL } \left(I, J\right) \right) \cup \left(A, J\right)\right \rbrace  \big \rbrace \\
%\nonumber
%\mbox{ if } I = \mbox{ First } \left(R, A \right) \wedge J < \mbox{ CloseL } \left(I, J \right)   \\
%\nonumber R , \text{ in any other case}
%\end{cases}
%\end{eqnarray}
%\end{definition}
\subsubsection{\label{subsubsec:modify}Modify}
This operation adds new information about an existing entity $A$ in the relation $R$. The modify operation does not remove any previous value of the object $A$. It closes the current version for the entity $A$ and adds a new version.

@@

\begin{definition}
 \emph{Modify(R, (A,I))}.
Consider the elements in definition \ref{def:valid-time-relation}. The algorithm for the modify operation is defined as follows.
\end{definition}

@@

\subsubsection{\label{subsubsec:insert}Insert}
The user wants to store an entity $A$ which is valid in the relation $R$ during the time interval specified by $I = \left(a, b \right)$.
%
%
%The interpretation of the insert operation is the following: The user wants to store in the database that an object $A$  is true for some period(s) of time $I$ in the given relation $R$. To indicate that the object $A$ is current in the relation, the special value Until Current, $\UC$  is used. 
There are two main cases when performing a create operation:
\begin{enumerate}
\item The entity $A$ was never in the relation $R$: The entity is added with the valid-time indicated by the crisp interval $I$. 
% In the example, this correspond with the following insertion sentences:
% \begin{verbatim}
% Insert(3,'Dracula',
% (15/3/2012, 30/3/2012));
% Insert(4,'Frankenstein',
% (25/3/2012, 4/4/2012));
% Insert(5,'Harry Potter',
% (18/3/2012, 2/4/2012));
% \end{verbatim}

\item The entity $A$ is in the relation $R$. Depending on the value of the time interval $I$, there are three possibilities:
	\begin{enumerate}
	\item Insert $\left(A, I\right)$ in the relation $R$. If the time interval $I$ does not overlap any other valid-time interval in the relation $R$ for the entity $A$. For example, consider that the book with ID = 3 is loaned on 4/4/2012 but the book has not been returned yet. This is illustrated in Table \ref{tbl:library-sample}.
%  The insert sentence is:
% 	      \begin{verbatim}
% Insert(3,'Dracula',
% (4/4/2012, UC));
% 	      \end{verbatim}

	\item Modify and close the current version of $A$ and insert $\left(A, I\right)$. For example, consider now that the patient with ID = 3 arrived to the hospital on 24/4/2012. Here the problem is that the patient with ID = 3 arrived on 4/4/2012 but, for some reason, the departure date was not stored. If the patient is already at the hospital, then it is necessary to set the departure date and add a new row with the new arrival date. This is illustrated in Table \ref{tbl:library-sample-2}.
%  the insert sentence is:
% 	    \begin{verbatim}
% Insert(3,'Dracula',
% (24/4/2012, UC));
% 	      \end{verbatim}
	\item Reject the insertion, if the time interval $I$ does overlap any existing valid-time interval for the entity $A$ in the relation $R$. For example, consider that the hospital manager wants to introduce a past arrival for the patient with ID = 3. The patient arrived on 6/4/2012 and left the hospital on 25/4/2012. As the dates do overlap, it is not possible that the patient was in the hospital at that time interval. Therefore, the insertion is rejected.
%  The insert sentence is:
% 
% 	      \begin{verbatim}
% Insert(3,'Dracula',
% (6/4/12/2012,25/4/2012));
% 	    \end{verbatim}

@@

\end{enumerate}
\begin{samepage}
\vglue13pt
%\begin{table}[htbp]
\tcap{\label{tbl:library-sample-2}Sample hospital database}
%\centerline{\small DATA TYPES}
\vglue-6pt
\centerline{\small\baselineskip=13pt
\begin{tabular}{c c c c}\\
\hline
\textbf{ID}  & Name  & Arrival & Leave \\
\hline
3 & John & 15/3/2012 & 30/3/2012 \\
4 & Antoon & 25/3/2012 & 4/4/2012 \\
5 & Daan Potter & 18/3/2012 & 2/4/2012\\
3 & John & 4/4/2012 & 23/4/2012 \\
3 & John & 24/4/2012 & UC \\
\hline\\
\end{tabular}
}
\end{samepage}

@@

\subsubsection{\label{subsubsec:del}Delete}
The delete operation logically removes a current entity $A$ which is valid in the relation $R$.
\begin{definition}
 \emph{Delete(r, t)}.
Consider the elements in definition \ref{def:valid-time-relation}.The algorithm for the delete operation is defined as follows.
% 
% Let $R \subseteq \A\  x\  \I$ be a valid-time relation and  $A=\left(A_1, \ldots, A_n \right), A \in \A$ be the set of attributes that define an entity and $I = \left(S_i,E_i\right)$ the set of attributes defining the time interval. The algorithm for the delete operation is defined as follows.
\end{definition}

@@

For example, consider that the hospital manager wants to delete the history for the patient with ID = 3. The result of this operation is shown in Table \ref{tbl:library-sample-3}.

@@

\begin{samepage}
\vglue13pt
%\begin{table}[htbp]
\tcap{\label{tbl:library-sample-3}Sample hospital database}
%\centerline{\small DATA TYPES}
\vglue-6pt
\centerline{\small\baselineskip=13pt
\begin{tabular}{c c c c}\\
\hline
\textbf{ID}  & Name  & Arrival & Leave\\
\hline
4 & Antoon & 25/3/2012 & 4/4/2012 \\
5 & Daan & 18/3/2012 & 2/4/2012\\
\hline\\
\end{tabular}
}
\end{samepage}

>>>>

\begin{itemize}
 \item A set of non-temporal attributes.
	\begin{equation}
	\label{eq:attribute-set}
	A = \left \lbrace A_1, A_2, \ldots, A_n \right \rbrace
	\end{equation}
      The domain for each attribute $A_1, \ldots, A_n$ is $D_1, \ldots, D_n$ respectively. 
\item The primary key $A_K$ is a subset of the attributes in $A$.
      \begin{equation}
       \label{eq:primary-key-a}
      A_K \subset A
      \end{equation}
\item A set of two attributes; $S$  and $E$ the attributes for the starting and ending points respectively. $I$ defines the valid time interval for the data. 
\begin{equation}
 \label{eq:attribute-time-interval}
I = \left \lbrace S, E \right \rbrace
\end{equation}
\item Then $R$, the schema for the valid-time relation is:
\begin{equation}
 \label{eq:valid-time-relation}
R = A \cup  I
\end{equation}
\item The primary key for the valid-time relation $R$ is:
\begin{equation}
 \label{eq:valid-time-temporal-pk}
PK = A_K \cup I
\end{equation}

@@

      \item Let $A_k \subset A$ be the set of the non-temporal attributes that are part of the primary key (equation \eqref{eq:primary-key-a}). Then, $a_k$ denotes the values for the non-temporal attributes of the primary key.
	    \begin{equation}
	     \label{eq:pk-attribute} 
	      a_k = t\left[A_K \right]
	    \end{equation}

@@

\begin{samepage}
\vglue13pt
%\begin{table}[htbp]
\tcap{\label{tbl:sample-definitions}Sample database containing the instance $r$ of the valid-time relation $R$.}
%\centerline{\small DATA TYPES}
\vglue-6pt
\centerline{\small\baselineskip=13pt
\begin{tabular}{c c c c c c}\\
\hline
& \textbf{A$_1$}  & \textbf{A$_2$}  & $A_3$ & $S$ & $E$ \\
\hline
$t_1$&$a_{11}$ & $a_{12}$ & $a_{13}$ & $s_1$ & $e_1$ \\
$t_2$ & $a_{21}$ & $a_{22}$ & $a_{23}$ &  $s_2$ & $e_2$ \\
$t_3$ & $a_{11}$ & $a_{12}$ & $a_{31}$ & $s_3$ & $e_3$\\
\hline\\
\end{tabular}
}
\end{samepage}

@@

For example, let us consider the last row in Table \ref{tbl:library-sample}. The value for the time interval is $I = \left(4/4/2012 , UC \right)$. The meaning is that the document with ID = 3 was valid the  4/4/2012 and it is still valid. The document with ID = 3 is current in the relation.

\begin{example}
\label{ex:library-database}
 Consider a historical database containing diplomatic documents. The starting and the ending dates when the diplomatic document is valid. It is possible that a diplomatic document is valid for a period of time and several years later it becomes valid again. The following elements are stored: an identifier of the document (ID), the entity that issues the document and the dates when document is valid. Table \ref{tbl:library-sample} illustrates the first version of the database, after three insertions. In this example, $A = ($ ID, Entity $)$ and $I = ($ Start, End $)$.
\end{example}

\begin{samepage}
\vglue13pt
%\begin{table}[htbp]
\tcap{\label{tbl:library-sample}Sample historical database}
%\centerline{\small DATA TYPES}
\vglue-6pt
\centerline{\small\baselineskip=13pt
\begin{tabular}{c c c c}\\
\hline
\textbf{ID}  & Entity  & Start & End \\
\hline
3 & E.U. & 15/3/2012 & 30/3/2012 \\
4 & N.A.T.O.  & 25/3/2012 & 4/4/2012 \\
5 & C.E.I. & 18/3/2012 & 2/4/2012\\
3 & E.U. & 4/4/2012 & UC \\
\hline\\
\end{tabular}
}
\end{samepage}

\begin{definition}
\label{def:current-in-relation}
\emph{Current $\left(r,a_k \right)$.}
Consider the elements in definition \ref{def:valid-time-relation}. The function Current $\left(r,a_k \right)$ returns the crisp time interval $i$ if the tuple $t$ with the primary key values given by $a_k$ is currently in the instance $r$ of the relation $R$. The function returns the empty set otherwise.
% 
% Consider $A=\left(A_1, \ldots, A_n \right), A \in \A$ be the set of attributes that define an entity and $I = \left(S_i, UC \right)$ the set of attribute defining the time interval. Let $R \subseteq \A\  x\  \I$ be a valid-time relation. The function Current $\left(R, A \right)$ returns the crisp time interval $I$ if the entity $A$ is currently in the relation $R$ or the empty set, if the entity $A$ is not currently in the relation. It is defined as follows:

@@

\begin{align}
\label{eq:current-in-relation}
\mbox{Current} \left(r, a_k\right) &=& \\ 
\begin{cases}
\nonumber
t & \mbox{ if } \exists t \in r : i = \left(s, UC \right) \mbox{ and }\\
& a_k = t[A_K] \\
\emptyset & \mbox{ otherwise}
\end{cases}
\end{align}
\end{definition}

For example, the function Current(r, (3,`E.U.')) returns the time interval $i = \left(4/4/2012 , UC  \right)$. Conversely, Current(r, (4,`N.A.T.O.')) returns the empty set.

@@

\begin{definition}
\label{def:close-current-version}
\emph{Close-current$\left(r, t\right)$}.
Consider the elements in definition \ref{def:valid-time-relation}. The function Close-current$\left(r, t \right)$ closes any current version $t_k$ of the entity given by $t$ if it exists and add the new version $t$. For the implementation the following variables are defined:
\begin{align}
 \label{eq:variables-current-version}
t_{CUR} = \mbox{ Current } \left(r, t[A_K] \right)\\
\nonumber
t_{UP} = t_{CUR}; t_{UP}\left[i \right] = \mbox{Close }\left(t_{UP}\left[i \right], t \right)
\nonumber
\end{align}
Then, $t_{CUR}$ is the current version of the entity given by the tuple $t$, and $t_{UP}$ is the updated version of the tuple $t_{CUR}$. In this updated version, the time interval given by $i_{UP}$ is closed with respect to the tuple $t$.

@@

For example, consider the document with ID = 3 in Table \ref{tbl:library-sample}. The current version of the document started on 4/4/2012. The document was not valid anymore but, for some reason, the date  was not registered. The document was valid again on 24/4/2012. Then, the Close-current function closes the current version of the document and adds a new version. First, the function CloseR is applied with $i_1 = \left(4/4/2012, UC \right)$ and $i_2 = \left(24/4/2012, UC \right)$. Hence, the value for the time interval $i_1$ is $ \left(4/4/2012, 23/4/2012 \right)$. Then, the modifications on the value for the time interval $i_1$ are stored. Finally a new row with the current values of the document and the time interval $i_2$ is stored.
% patient with ID = 3 in Table \ref{tbl:library-sample}. The current version arrived on 4/4/2012. The patient left the hospital but, for some reason, the date  was not registered. The patient arrived again at the hospital on 24/4/2012. Then, the Close-current function allows to close the current version of the patient and insert a new version. First, the function CloseR is applied with $I = \left(4/4/2012, UC \right)$ and $J = \left(24/4/2012, UC \right)$. Hence, the value for the time interval $I$ is $ \left(4/4/2012, 23/4/2012 \right)$. Then, the modifications on the value for the time interval $I$ are stored. Finally, a new row with the current values of the patient and the time interval $J$ is stored.
The result of this operation is illustrated on Table \ref{tbl:library-sample-2}.

%It is also possible to append at the beggining by using \eqref{def:close-a-crisp-interval-l} and \eqref{def:first-in-relation}.
%\begin{definition}
%\label{def:append-first-version}
%Consider an object $A$, the relation $R$ and a time interval $J$. Then, the function append-first$\left(R,A,J \right)$ closes any right open version of the object $A$ if it exists and add the new version $\left(A, J \right)$.
%
%\begin{eqnarray}
%%\label{eq:close-current}
%\text{Append-first} \left(R, A, J \right) =\\
%\begin{cases}
%\nonumber
%\big \lbrace R - \left(A, I \right) \cup \left \lbrace \left(A, \mbox{ CloseL } \left(I, J\right) \right) \cup \left(A, J\right)\right \rbrace  \big \rbrace \\
%\nonumber
%\mbox{ if } I = \mbox{ First } \left(R, A \right) \wedge J < \mbox{ CloseL } \left(I, J \right)   \\
%\nonumber R , \text{ in any other case}
%\end{cases}
%\end{eqnarray}
%\end{definition}
\subsubsection{\label{subsubsec:modify}Modify}
This operation adds new information about an existing entity (given by the tuple $t$) in the instance $r$ of the relation $R$. The modify operation does not remove any previous value of the entity. It closes the current version and adds a new version.

@@

\begin{definition}
 \emph{Modify(r, t)}.
Consider the elements in definition \ref{def:valid-time-relation}. The algorithm for the modify operation is defined as follows.
\end{definition}

@@

\subsubsection{\label{subsubsec:insert}Insert}
The user wants to store an entity (given by the tuple $t$) which is valid in the instance $r$ of the relation $R$ during the time interval specified by $i = \left(s, e \right)$.
%
%
%The interpretation of the insert operation is the following: The user wants to store in the database that an object $A$  is true for some period(s) of time $I$ in the given relation $R$. To indicate that the object $A$ is current in the relation, the special value Until Current, $\UC$  is used. 
There are two main cases when performing a create operation:
\begin{enumerate}
\item The entity was never in the relation: The entity is added with the valid-time indicated by the crisp interval $i$. 
% In the example, this correspond with the following insertion sentences:
% \begin{verbatim}
% Insert(3,'Dracula',
% (15/3/2012, 30/3/2012));
% Insert(4,'Frankenstein',
% (25/3/2012, 4/4/2012));
% Insert(5,'Harry Potter',
% (18/3/2012, 2/4/2012));
% \end{verbatim}

\item The entity is in the relation. Depending on the value of the time interval, there are three possibilities:
	\begin{enumerate}
	\item Insert $t$ in the instance $r$ of the relation $R$. If the time interval $i$ does not overlap any other valid-time interval in the instance $r$ relation $R$ for the entity. For example, consider that the document with ID = 3 began again to be valid on 4/4/2012 and it is still currently valid. This is illustrated in Table \ref{tbl:library-sample}.
%  The insert sentence is:
% 	      \begin{verbatim}
% Insert(3,'Dracula',
% (4/4/2012, UC));
% 	      \end{verbatim}

	\item Modify and close the current version of $t$ and insert the new version. For example, consider now that the document with ID = 3 was valid on 24/4/2012. Here the problem is that the document with ID = 3 was valid on 4/4/2012 but, for some reason, the ending date was not stored. If the document is again valid, then it is necessary to set the ending date and add a new row with the new starting date. This is illustrated in Table \ref{tbl:library-sample-2}.
%  the insert sentence is:
% 	    \begin{verbatim}
% Insert(3,'Dracula',
% (24/4/2012, UC));
% 	      \end{verbatim}
	\item Reject the insertion, if the time interval $i$ does overlap any existing valid-time interval for the entity $t$ in the instance $r$ of the relation $R$. For example, consider that the document manager wants to introduce a past valid-time for the document with ID = 3. The validity period for the document is $[$6/4/2012, 25/4/2012$]$. As the dates do overlap, it is not possible that the document was valid at that time interval. Therefore, the insertion is rejected.
%  The insert sentence is:
% 
% 	      \begin{verbatim}
% Insert(3,'Dracula',
% (6/4/12/2012,25/4/2012));
% 	    \end{verbatim}

@@

\end{enumerate}
\begin{samepage}
\vglue13pt
%\begin{table}[htbp]
\tcap{\label{tbl:library-sample-2}Sample historical database}
%\centerline{\small DATA TYPES}
\vglue-6pt
\centerline{\small\baselineskip=13pt
\begin{tabular}{c c c c}\\
\hline
\textbf{ID}  & Entity  & Start & End \\
\hline
3 & E.U. & 15/3/2012 & 30/3/2012 \\
4 & N.A.T.O. & 25/3/2012 & 4/4/2012 \\
5 & C.E.I.  & 18/3/2012 & 2/4/2012\\
3 & E.U. & 4/4/2012 & 23/4/2012 \\
3 & E.U. & 24/4/2012 & UC \\
\hline\\
\end{tabular}
}
\end{samepage}

@@

\subsubsection{\label{subsubsec:del}Delete}
The delete operation logically removes a current entity $t$ which is valid in the instance $r$ of the relation $R$.
\begin{definition}
 \emph{Delete(r, t)}.
Consider the elements in definition \ref{def:valid-time-relation}.The algorithm for the delete operation is defined as follows.
% 
% Let $R \subseteq \A\  x\  \I$ be a valid-time relation and  $A=\left(A_1, \ldots, A_n \right), A \in \A$ be the set of attributes that define an entity and $I = \left(S_i,E_i\right)$ the set of attributes defining the time interval. The algorithm for the delete operation is defined as follows.
\end{definition}

@@

For example, consider that the document manager wants to delete the history for the document with ID = 3. The result of this operation is shown in Table \ref{tbl:library-sample-3}.

@@

\begin{samepage}
\vglue13pt
%\begin{table}[htbp]
\tcap{\label{tbl:library-sample-3}Sample historical database}
%\centerline{\small DATA TYPES}
\vglue-6pt
\centerline{\small\baselineskip=13pt
\begin{tabular}{c c c c}\\
\hline
\textbf{ID}  & Entity  & Start & End\\
\hline
4 & N.A.T.O. & 25/3/2012 & 4/4/2012 \\
5 & C.E.I. & 18/3/2012 & 2/4/2012\\
\hline\\
\end{tabular}
}
\end{samepage}
